---
title: "Analysing electricity usage to check with rate is best to go with"
author: "Andrew Edwards"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = TRUE,
  cache_path = "hydro-calc-cache/",
  fig.path = "hydro-calc-fig-cache/",
  fig.width = 7,
  fig.height = 6
)
  # comment = "#>",
```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("hydro-calc.Rmd")
```

```{r, packages, echo = FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
# load_all()
```

# Load data

```{r, load}
raw <- read.csv("bchydro.com-consumption-XXXXXXXX0569-2025-09-02-200019.csv",
                header = TRUE,
                stringsAsFactors = TRUE) %>%
  as_tibble()

summary(raw)
```

Shows that we only need to keep a few columns (and two look to be the same)
```{r, refine}
raw_reduced <- select(raw,
                      Interval.Start.Date.Time,
                      Inflow..kWh.,
                      Net.Consumption..kWh.)
summary(raw_reduced)

expect_equal(raw_reduced$Inflow..kWh.,
             raw_reduced$Net.Consumption..kWh.)

dat <- select(raw_reduced,
              time = Interval.Start.Date.Time,
              use = Inflow..kWh.) %>%
  mutate(time = lubridate::ymd_hm(time))

summary(dat)
```

Looked earlier like there were three repeated times:
```{r, repeated}
repeated <- summary(as.factor(dat$time))
rep <- repeated[repeated > 1]

rep
filter(dat,
       time == names(rep[1]))
```
Not quite working, don't worry as only three times.

Add separate date and hour columns, to then summarise on
```{r, expand}
dat_expand <- mutate(dat,
                     date = date(time),
                     year = year(time),
                     month = month(time),
                     week = week(time),
                     hour = hour(time))
dat_expand

summary(dat_expand)

# Probably when heat pump and then car charger installed (and something else?):
filter(dat_expand,
       use == 0)
```

TODO check start and end, exclude any non-full days.


Aggregate daily, and some key dates
```{r, daily}
heat_pump_start <- ymd("2023-06-21")
ev_start <- ymd("2024-10-12")
dat_daily <- summarise(group_by(dat_expand,
                                date),
                       daily = sum(use))
dat_daily

plot(dat_daily,
     type = "l",
     xlab = "Date",
     ylab = "Daily usage (kWh)",
     ylim = c(0, 130))
abline(v = heat_pump_start,
       col = "grey")
abline(v = ev_start,
       col = "grey")
text(heat_pump_start,
     130,
     "Heat pump",
     pos = 4)
text(ev_start,
     130,
     "Car charger",
     pos = 4)

```
Presume those spikes are restarting the hot tub (could check).

## Plot usage by day of year, colour for each year

Adapting from `plot.pacea_buoy()`, but a bit busy, so going to then do by week
and then by month.

```{r dailyplot}
g <- ggplot(dat_daily,
            aes(lubridate::yday(date),
                daily,
                group = factor(lubridate::year(date)),
                colour = factor(lubridate::year(date))))

h <- g + geom_line(na.rm = TRUE) +
  scale_y_continuous(limits = c(0, max(dat_daily$daily)),
                     expand = c(0, 10)) +
                       # ylim(0, max(dat_daily$daily)) +
  theme_bw() +
#  theme(plot.title = element_text(hjust = 0.5),
#        strip.background = element_rect(colour = "grey70",
#                                        fill = "grey95")) +
  labs(x = "Julian Day",
       colour = "Year",
       title = "Usage by day") +
  scale_colour_discrete() +
  # scale_colour_viridis_c() +
  ylab("Daily usage (kWh)")
h
```

### Weekly

```{r, weekly}
dat_weekly <- summarise(group_by(dat_expand,
                                 week,
                                 year),
                        weekly = sum(use),
                        num_days = length(unique(date))) %>%
  mutate(weekly_scaled = weekly * 7 / num_days)

dat_weekly

table(dat_weekly$num_days)

dat_weekly %>% as.data.frame()

```
Now plot
```{r weeklyplot}
g <- ggplot(dat_weekly,
            aes(week,
                weekly_scaled,
                group = factor(year),
                colour = factor(year)))

h <- g + geom_line(na.rm = TRUE,
                   linewidth = 2) +
  scale_y_continuous(limits = c(0, max(dat_weekly$weekly_scaled)),
                     expand = c(0, 10)) +
  theme_bw() +
  labs(x = "Week of year",
       colour = "Year",
       title = "Usage by week") +
  scale_colour_discrete() +
  # scale_colour_viridis_c() +
  ylab("Weekly usage (kWh)")

h
```





TODO:

Add points on representing start dates as on full plot.

Plot the 11 months that we've had the car, compared to the same dates for the
previous 11 months. Actually just sum the total for each first of all.
xccCan assume somewhat that the difference is the car use. Work
that out, and then see savings for having that at night.

Did a rough calc of $300 anyway. Be good to do each of the four hydro options.
